---
title: "ontoProc: Ontology interfaces for Bioconductor, with focus on cell type identification"
author: "Vincent J. Carey, stvjc at channing.harvard.edu"
date: "`r format(Sys.time(), '%B %d, %Y')`"
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{ontoProc: some ontology-oriented utilites with single-cell focus for Bioconductor}
  %\VignetteEncoding{UTF-8}
bibliography: ontobib.bib
output:
  BiocStyle::html_document:
    highlight: pygments
    number_sections: yes
    theme: united
    toc: yes
---



```{r}
suppressPackageStartupMessages({
library(knitr)
library(ontoProc)
library(ontoProc2)
cl = retrieve_semsql_conn(ontology = "cl")
cl = semsql_to_oi(con= cl)
go = retrieve_semsql_conn(ontology = "go")
#takes a few minutes
go = semsql_to_oi(con= go)
library(ontologyPlot)
library(BiocStyle)
})
```
 The protein onotology is large and takes 5 minutes to download using wget on the terminal . To take the protein ontology to and object take over a hour. Due to time I have a example saved to for faster processing . Below is the code that  to put in a R script and run in the background.
 The terminal command that was uses is following:
 
Rscript pronto.R  1> pronto.log 2> pronto.err &
```{r ,eval=FALSE}
#the code in the pronto.R
library(ontoProc2)
library(ontologyIndex)
pr = retrieve_semsql_conn(ontology = "pr")
pr = semsql_to_oi(con = pr)

saveRDS(pr, file = "pr.rds", compress ='xz')
```

```{r}
pr = readRDS("~/pr.rds")
#pr = readRDS(system.file("extdata/pr.rds", package="ontoProc2"))
```


```{r lklk}
kable(stab <- seur3kTab())
```




```{r}
onto_plot2(cl, stab$tag)
```





```{r lksig}
sigels = c("CL:X01"="GRIK3", "CL:X02"="NTNG1", "CL:X03"="BAGE2",
             "CL:X04"="MC4R", "CL:X05"="PAX6", "CL:X06"="TSPAN12", 
             "CL:X07"="hSHISA8", "CL:X08"="SNCG", "CL:X09"="ARHGEF28", 
             "CL:X10"="EGF")
```

### A data.frame defining the cell types and their properties

The `cyclicSigset` function produces a data.frame instance
connecting cell types with the genes expressed or unexpressed.
```{r lkdfff}
cs = cyclicSigset(sigels)
dim(cs)
cs[c(1:5,9:13),]
table(cs$cond)
```

# Subsetting SingleR resources using ontological mapping

### A data.frame mapping from informal to formal terms

Aaron Lun has produced a mapping from informal terms used in the
Human Primary Cell Atlas to Cell Ontology tags.  We provisionally
include a copy of this mapping in ontoProc:

```{r lkmap}
hpca_map = read.csv(system.file("extdata/hpca.csv", package="ontoProc"), strings=FALSE)
head(hpca_map)
```

We will rename columns of this map for convenience of our `bind_formal_tags` method.
```{r doren}
names(hpca_map) = c("informal", "formal")  # obligatory for now
```

### Binding formal tags to the HPCA data

I am turning this code off for now because there is no
standard approach to getting the mapping from the SummarizedExperment
yet.  When SingleR merges the 'standardized' branch, this will come back.


let's retrieve the HPCA data from SingleR:
```{r gethpca, eval=TRUE}
library(SingleCellExperiment)
library(celldex)
hpca_sce = HumanPrimaryCellAtlasData()
```
Now bind the formal tags:
```{r dobind, eval=TRUE}
hpca_sce = bind_formal_tags(hpca_sce, "label.fine", hpca_map)
length(unique(hpca_sce$label.ont))
```
We don't check for failed mappings:
```{r justna, eval=TRUE}
length(xx <- which(is.na(hpca_sce$label.ont)))
if (length(xx)>0) print(colData(hpca_sce)[xx,])
sum(hpca_sce$label.ont == "", na.rm=TRUE) # iPS and BM
```

### Subsetting using the class hierarchy of Cell Ontology

```{r dosub, eval=TRUE}
hpca_mono = subset_descendants( hpca_sce, cl, "^monocyte$" )
table(hpca_mono$label.fine)
```

```{r}
table(hpca_mono$label.ont) # not much diversity
```


```{r}
hpca_tcell = subset_descendants( hpca_sce, cl, "^T cell$" )
table(hpca_tcell$label.fine)
```

```{r}
table(hpca_tcell$label.ont) 
```

```{r}
uu = unique(hpca_tcell$label.ont)
onto_plot2(cl, uu)
```

```{r}

dbDisconnect()


```

